### 2019-10-22 ms/ms.tex
Track stats; intro

\documentclass[12pt]{report}
\usepackage[utf8]{inputenc}
\usepackage{times}
\renewcommand\thesection{\arabic{section}}
\usepackage[parfill]{parskip}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage[left=0.75in,right=0.75in,top=0.75in,bottom=0.75in]{geometry}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{nth}
\usepackage{bm}
\usepackage{siunitx}
\usepackage[labelfont=bf]{caption}
\usepackage[doublespacing]{setspace}
\usepackage{filecontents}
\usepackage{lineno} \linenumbers
\usepackage{amsmath}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Bibliography%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[style=numeric-comp,sorting=none,uniquelist=minyear,uniquename=false,maxcitenames=2,maxbibnames=5,doi=false,url=false,isbn=false,defernumbers=true,giveninits=true]{biblatex}
%Italicise et al.
\renewbibmacro*{name:andothers}{
% Based on name:andothers from biblatex.def
\ifboolexpr{ test{\ifnumequal{\value{listcount}}{\value{liststop}}} and test \ifmorenames } {\ifnumgreater{\value{liststop}}{1} {\finalandcomma}
{}%
\andothersdelim\bibstring[\emph]{andothers}} {}}
%Make 'and' an ampersand
\renewcommand*{\finalnamedelim}{%
\ifnumgreater{\value{liststop}}{2}{\finalandcomma}{}%
\addspace\&\space}
%Remove 'in:'
\renewbibmacro{in:}{}
%Remove pp. on pages numbers
\DeclareFieldFormat[article]{pages}{#1}%
%Remove unwanted entries
\AtEveryBibitem{%
  \clearfield{day}%
  \clearfield{month}%
  \clearfield{endday}%
  \clearfield{endmonth}%
  \clearfield{url}%
  \clearfield{URL}%
  \clearfield{eprint}%
}
%Remove spaces between initials
%Remove double quotes around titles
\DeclareFieldFormat*{citetitle}{#1}
\DeclareFieldFormat*{title}{#1}
\addbibresource{/home/will/Papers/Library.bib}
\begin{document}
\section*{Title page}
\textbf{Article title}: The relative velocity of climate change
reveals remarkable idiosyncrasy of species' distributional shifts

\textbf{Authors:} William D.\ Pearse$^{1,*}$ \& T.\ Jonathan
Davies$^{2,*}$

$^1$ Department of Biology \& Ecology Center, Utah State University,
5305 Old Main Hill, Logan UT, 84322

$^2$ Department of Biological Sciences, Universit\'{e} du Qu\'{e}bec
\`{a} Montr\'{e}al, Montr\'{e}al, Qu\'{e}bec, Canada

$^*$ To whom correspondence should be addressed:
\url{will.pearse@usu.edu} and \url{j.davies@ubc.ca}

\textbf{Word-count}: XXX (excluding figure legends)

\clearpage
\textbf{To date, our understanding of how species have shifted in
  response to recent climate warming has been based on a few studies
  with limited number of species. Here we present a comprehensive,
  global overview of species' distributional responses to changing
  climate across a broad variety of taxa (animals, plants, and
  fungi). We characterise species' responses using a metric that
  describes the biotic velocity of climate change: how closely
  species' responses have tracked changing climate through time
  climate. In contrast to existing metrics that focus purely on space,
  we focus on species, and thus can examine drivers of interspecific
  variation. Here we show that species responses are highly
  idiosyncratic, and thus previous analyses assuming uniform responses
  may be misleading. This is in stark contrast to species' static
  present-day and past distributions, which show strong evidence of
  the imprint of evolutionary history and functional traits. Our
  analyses are a first step in exploring the vast wealth of empirical
  data on species' historic responses to recent climate change. We
  highlight that our findings, while surprising, are consistent with
  paleontological patterns observed at the onset of mass extinction
  events.}
% Is that last sentence a bit much?

The natural environment is changing rapidly. Over the last century,
habitats have become increasingly fragmented and
isolated\supercite{Haddad2015}, the climate has become warmer, and
extreme climate events more frequent\supercite{IPCC2014}. Biological
diversity---the species of plants and animals with which we cohabit
the Earth---has responded predictably. Present-day extinction rates
are estimated to be three orders of magnitude greater than background
rates \supercite{Pimm2014} and are projected to increase further over
the next several decades \supercite{MEA2005}. Current estimates
suggest that over at least one million species may be threatened with
extinction\supercite{IPBESland2019}, and that we have experienced a
50\% decline in animal diversity over the past 40 years
\supercite{McLellan2014}.

The main direct human-induced drivers that impact biodiversity now are
habitat loss \supercite{Newbold2015} and fragmentation
\supercite{Haddad2015}, but climate change is likely to become a
dominant driver in the next few
decades\supercite{Parmesan2003,Thuiller2007,Urban2015}. Projected
impacts of climate change on biodiversity have attracted much
attention, and it has been suggested that up to 37\% of species may be
committed to global extinction by 2050\supercite{Thomas2004}. In the
face of changing climate species must either shift in space, to track
favourable conditions, or time, for example, flowering and breeding
earlier, or migrating sooner. Evidence suggests they are doing both
\supercite{Parmesan2003,Parmesan2006,Menzel2006,Chen2011}, but not all
species are responding equally. Some species are expanding their
distributions, sometimes invasively \supercite{Hellmann2008}, and some
species are flowering later\supercite{Cook2012}; species may also be
simply becoming less predictable \supercite{Pearse2017phenology}.
% I feel this isn't just a token cite; we want to set up our result a
% little, I think
Developing robust predictions of ecological responses to climate
change thus requires data from diverse species, ecosystems, and
climates to account for these varying drivers. By contrast, current
projections frequently assume a simplistic or uniform response, and
have tended to overlook interspecific variation.
% I'm not sure I necessarily agree with this statement

We currently lack robust estimates of how species are shifting their
distributions in response to recent warming. The velocity of
temperature change---the rate a species would have to shift its range
to maintain constant temperature---has been calculated at 0.42 km/yr
\supercite{Loarie2009}. A meta-analysis of taxonomic groupsestimated
the average pole-ward shift in species' distributions has been
estimated as approximately 1.69 km/yr-1 \supercite{Chen2011},
suggesting that species might be capable of keeping pace with shifting
climate. Yet this estimate is derived from just a handful of studies,
and does not address the large variation in response among species.
Here we leverage the vast volume of distributional data stored in
digital data repositories\supercite{url_gbif} to provide the first
global synthesis of species distributional responses to climate
change. By contrasting historical niche dimensions with those of the
present day, we can characterise the true biotic velocity of climate
change, which represents the actual velocity with which species track
climate across space.

We show that, on average, species' historical (pre-1970) ranges have
warmed significantly, and that species have generally shifted their
distributions to ameliorate this warming. Yet this general pattern
masks a remarkably degree of variability in species responses: many
species' ranges are now even warmer than if their distibutions had
remained static. Further, we find that, while species' distributions
are strongly predicted by their evolutionary history and functional
traits, the same is not true of their degree of tracking. Thus while
species' static distributions are reasonably predictable, their
changes in distributions are not.

\clearpage
\subsection*{Results and discussion}
\textbf{A comparative index of species' niche tracking.} We propose to
measure the extent to which a species' distribution is tracking
climate change by standardising its observed change by the change it
would have experienced had it not shifted its distribution. Formally,
we define this as:

\begin{equation}
  track_\tau = \frac{present_\tau - past_\tau}{projected_\tau - past_\tau}
\end{equation}

Where $current_\tau$ is a given quantile ($\tau$) of a species'
climatic distribution across its present-day range, $past_\tau$ the
same quantile of a species' range in the past, and $projected_\tau$
the same quantile of a species' present-day range under past climatic
conditions. This index should be calculated separately for
temperature, pollution, or whatever environmental variable is of
interest. Thus a $track$ of 0 would indicate a species that has
tracked changing climate perfectly, while a $track$ of 1 would be
consistent with a species that had not moved whatsoever to track
changing environmental conditions. Values between 0 and 1 suggest
moderate tracking, while values greater than 1 or less than 0
represent a change in climate relationship greater than if the species
had not moved at all. Uncertainty in either species' past or present
ranges can be accounted for through a bootstrap procedure, whereby the
past and/or present distributions are re-sampled (with replacement) to
match the sampling of the other time-period, and the standard
deviation of the index across those samples reported (see methods).

Our approach has two main innovations: its focus on (1) the whole of a
species' distribution and(2) expected change. (1) We suggest that
something of a disconnect is developing among related fields of
ecology: species' range modelers typically focus on species'
limits\supercite{Sexton2009} (\emph{i.e.}, the upper and lower
quantiles) while community ecologists \supercite{Tilman2004} and
macro-evolutionary \supercite{Zanne2018} biologists focus on central
tendencies (\emph{i.e.}, the median). We therefore see no strong
\emph{a priori} reason that species can, or must, track all quantiles
equally, and so our approach is to check all quantiles of a species'
distribution. (2) The denominator of our index re-scales species'
change around an expectation under distributional stasis. We suggest
that species' changes should be measured on the basis of the magnitude
of change to which they are exposed. This change reflects the
magnitude of ecological and evolutionary selection, and while we
suggest there are many ways of addressing this, we suggest ours is the
simplest and so perhaps a fair starting point. We note that our index
could by re-scaling according to the distance that would need to be
travelled to match the degree of change (\emph{sensu}
\citeauthor{Loarie2009}\supercite{Loarie2009}), but prefer to present
here a dimensionless (and so unitless) index of tracking.
% ...something like this? I re-wrote everything to be about the
% "relative velocity of change" (which would have given us some fancy
% notation if we wanted it) but as we don't use distances or speeds I
% feel we ought not to?

\textbf{Niche tracking across the tree of life throughout the past
  sixty years.} To assess the extent to which species have tracked
climate, we used the most comprehensively sampled time-series data for
species' distributions (GBIF\supercite{url_gbif}) and climate (UEA's
CRU\supercite{Harris2014}). We focus on five quantiles: the \nth{5},
\nth{25}, \nth{50} (median), \nth{75}, and \nth{95}. In the methods we
give more detail of our reproducible analysis pipeline, and in the
supplementary materials we repeat our analyses with different data
curation and cleaning choices. Here, however, we present species'
range data collected from 1955--2015 for 10,700 species (4,879 plants,
1,517 fungi, 273 mammals, 433 birds, 147 reptiles and amphibians, and
3,451 insects), each observation matched to one of nine climate
variables in the year in which they were sampled (mean, minimum,
maximum temperature, diurnal temperature range, precipitation, wet-day
frequency, vapour pressure, and cloud cover). Our metric was designed
to account for known biases in the GBIF data \supercite{Beck2014}, and
we present in the supplementary materials additional analyses that
address and account for such biases.

Figure \ref{violin} shows the empirical index values for our focal
taxa for mean temperature. Clear patterns emerge: 99.5\% of species
are tracking some climate variable through time, and a large number of
species are, to some extent, tracking temperature through time (65\%
track at least one quantile, 43\% track at least two, and 4\% track
all five). The three temperature indices are the least-tracked of any
climate variable, while precipitation is the most-tracked (86\% of
species track at least one of its quantiles). Notably, however, many
taxa are not tracking climate through time, and many have experienced
more change than if their distributions had remained static (99.9\% of
species have `overshot' in at least three quantiles). As we show in
figure \ref{violin}, and explore further in the supplementary
materials, these results are consistent with our null simulations
where species' range change is a combination of essentially random
movement and a moderate degree of subsequent climate filtering.

\textbf{Idiosyncratic drivers of species' tracking.} To examine the
potential drivers of species' ranges and range tracking, we first
examined the phylogenetic signal of species' climate quantiles. We
focused on mammals\supercite{xxx}, birds\supercite{xxx},
amphibians\supercite{xxx}, reptiles\supercite{xxx}, and
plants\supercite{xxx}, since we were able to find broadly inclusive,
dated phylogenies for these taxa. We found statistically significant
evidence of phylogenetic signal across XXX of 135
taxon-climate-quantile combinations of species' past ranges (and XXX
in the present; see figure \ref{phylogeny}). This is consistent with
the known patterns of phylogenetic conservatism outlined above. For
our index, however, we found evidence of phylogenetic signal in only
XXX combinations (XXXlist themXXX). This confirms the known results
that species' present-day and past ranges are strongly influenced by
evolutionary, but suggests that the magnitude of change (relative to
expectations) may be driven by different drivers.

We additionally examined species' traits as potential drivers of
species' distributions and potential to track changing environment. We
focused on mammals\supercite{xxx}, birds\supercite{xxx}, and
plants\supercite{xxx}, since global, openly-released trait datasets
exist for these taxa. Again, we found statistically significant
associations between traits and past distributions of XXX of XXX
taxon-climate-quantile combinations (XXX in the present; see figure
\ref{traits}). While the traits we used are strongly associated with
species' climate responses\supercite{xxx} (hence their strong
predictive power of species' distributions, but not tracking index),
we emphasize that we do not examine traits related to dispersal
ability. This is due to a paucity of open-access dispersal traits
\supercite{xxx}, and we highlight that future work should address
whether our index is explained by such drivers.
% Another potential problem here is no trait*env interaction, which I
% sense is also important...
% WILL - re-do TEMP.INDEX ~ TEMP! THERMOPHILES!!!

\textbf{Climate space is not the same as tracking space.} Given our
index is poorly predicted by traditionally important variables, we now
consider whether index values are predictive of each other. It is well
known that global climate\supercite{xxx}, and so species' positions
across the various axes of climate space\supercite{xxx}, are not
independent, and so we set out to test whether species' tracking
indices are also independent. As we show in figure \ref{pca}, a
principal components analysis of species' present, past, and projected
climate distributions shows strong correlations across climate
variables. Yet the index does not: species' relative degree of
tracking is both of much higher dimension, and shows relatively less
correlation across climates axes. We argue, therefore, that while
species' positions in climate space show strong associations, species'
relative tracking of their position in climate space does not.

If true, this result would explain the lack of association between
species' traits and evolutionary history with their degree of climate
tracking. The ecological and evolutionary rules that determine
species' climate relationships are changing, and thus the factors that
once are strong predictors of past and present distributions are not
predictive of the differences between them. Yet there are alternative
explanations, of which we highlight three: (1) that we have missed
fundamentally important environmental axes, (2) biotic interactions,
and (3) that our index highlights random variation in
distribution. (1) We fully acknowledge that we ignore potentially
relevant environmental drivers such as geology (and so soil), habitat,
and all Anthropogenic factors other than climate. We are not alone in
this simplification\supercite{xxx}, but we detect strong evolutionary
pattern in the factors we do examine in the data underlying our
index. (2) Competition and facilitation among species are likely
strong drivers of species' ability to track climate\supercite{xxx}. We
acknowledge it is entirely possible that ignoring these factors is the
driver of our result and, if so, we urge for an acceleration of the
inclusion of such information into broad-scale analyses
\supercite{xxx}. (3) If species were perfectly tracking climate
(\emph{i.e.}, $present_\tau$ is equal to $past_\tau$ in the absence of
sampling error or some other stochastic process) then our index values
would essentially be random numbers. There are good empirical evidence
that this is not the case \supercite{xxx} but, as we show in the
supplementary materials, simulations suggest out values are
incompatible with this result.

\subsection*{Conclusion}
Our results are consistent with the processes underlying species'
responses to climate change being distinct from those that generated
their ranges in the past. That is not to say that we do not detect the
imprint of ecological and evolutionary history, or that those earlier
processes are of predictive power for species distributions, but
rather that to understand change we must consider new drivers. We do
not know what those drivers are, although we highlight that biotic
interactions may play a part. We also highlight that our data,
although comprehensive, are not taxonomically complete (we address
only XXX species) and are necessarily spatially ($0.5^\circ$ grid
cells) and temporally (yearly) coarse given the extent (global and 60
years) of our study. We are not the first to suggest that, for large
numbers of species, the extent of climate tracking is idiosyncratic
\supercite{xxx}. These results are, however, consistent with the
paleo-ecological record. During and after the last ice-age,
`no-analogue' communities generated previously-unseen species
assemblages \supercite{xxx}. Once previous mass extinction events were
underway, species' extinctions became unpredictable
\supercite{xxx}. Scientists have been warning for decades that we are
beginning a sixth mass extinction;
% Soule 1999 but find something earlier
perhaps we are seeing the end of its beginning.

This consensus of species' past and present ranges broadly tracking
the same underlying climate optima is complicated by paleo-ecological
studies that have revealed greater variability of species' historic
ranges in previous centuries and millennia than previously thought
\supercite{Williams2007,Veloz2012,Maguire2015}.

\clearpage
% Figures

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=0.7\textwidth]{../figures/violin-tmp-50.pdf}
  \end{center}
  \caption{\doublespacing \textbf{Empirical distributions of track
      index for \nth{50} of mean temperature.} Need to add simulations
    to plot. Add phylo-pics?}
  \label{violin}
\end{figure}

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=0.7\textwidth]{../figures/pca-50.pdf}
  \end{center}
  \caption{\doublespacing \textbf{Principal components analyses of
      present and past climate, along with our index of climate
      tracking.} Loadings are correlated (eigenvalues shown at bottom
    of plot), but much weaker overall in index.}
  \label{pca}
\end{figure}

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=0.7\textwidth]{../figures/phylo-tmp-50.pdf}
  \end{center}
  \caption{\doublespacing \textbf{Phylogeny with ancestral states
      branches of past temperature (50th quantile) and tips of the
      temperature 50th index.} Scale between the two kinds of input
    data currently is slightly off, hence no scale bar, and obviously
    the tip colours are hideous, but I think no one would disagree
    there's signal in the underlying data!}
  \label{pca}
\end{figure}

% zoom in on one clade and show that the tips are a random subset (but the branches are not)
% Will wonders if we can put the correlation matrix to the side of an upward-facing tree
% have lines going outward from the tree to show mapping

\begin{figure}[h!]
  \begin{center}
  \end{center}
  \caption{\doublespacing \textbf{Trait correlation box.}
    XXXmissingXXX}
  \label{traits}
\end{figure}

\clearpage

\textbf{Acknowledgments} WDP was funded by NSF ABI-1759965, NSF
EF-1802605, and USDA Forest Service agreement 18-CS-11046000-041. TJD
was funded by \emph{Fonds de Recherche Nature et Technologies} grant
number 168004.

\textbf{Author Contributions} WDP and TJD contributed to all aspects
of the study.

\textbf{Author information} Reprints and permissions information is
available at \url{www.nature.com/reprints}. The authors declare no
competing financial interests. Correspondence and requests for
materials should be addressed to \url{will.pearse@usu.edu}.

\textbf{Competing interests} The authors declare no competing
financial interests

\clearpage

\section*{\Large Methods}

\printbibliography

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: